

```{r}
setwd(here::here('40_Aug_09_2023'))
library(tidyverse)
```




```{r}
set.seed(2345345)
N <- 100
generated_data <- tibble(
  year = rep(seq(1950, 2020, 10), each = N),
  category = sample(
    paste0('Cat.', 1:4), 
    prob = c(1 / 8, 1 / 2, 1 / 4, 1/8), 
    size = length(year), 
    replace = TRUE
  ) 
) |> 
  count(year, category)
generated_data
# # A tibble: 32 × 3
#     year category     n
#    <dbl> <chr>    <int>
#  1  1950 Cat.1       14
#  2  1950 Cat.2       53
#  3  1950 Cat.3       25
#  4  1950 Cat.4        8
#  5  1960 Cat.1       17
#  6  1960 Cat.2       37
#  7  1960 Cat.3       26
#  8  1960 Cat.4       20
#  9  1970 Cat.1       18
# 10  1970 Cat.2       51
```


```{r}
standard_heatmap <- generated_data |> 
  ggplot(aes(year, category, fill = n)) +
  geom_tile(color = 'white') +
  labs(
    x = element_blank(), 
    y = element_blank(), 
    fill = 'Count',
    title = 'This is a fabricated heat map',
    subtitle = 'Let\'s make this look nice'
  ) +
  theme_minimal(base_size = 20, base_family = 'Source Sans Pro')
```



Now let's move the guide and titles.

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    legend.position = 'top'
  )
```


Make the color bar nicer

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    legend.position = 'top'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  )
```


Move the color bar to a better position.
The moment you to that you need to set `legend.direction` and make some room for the legend (e.g. by setting margins on `plot.subtitle`)

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.4, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  )
```


Get rid of extra spacing (readjust legend position)

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.355, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  ) +
  coord_cartesian(expand = FALSE)
```


Choose a new color
Continuous scale that makes sense in this case. 

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.355, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  ) +
  coord_cartesian(expand = FALSE) +
  scale_fill_distiller(type = 'seq', palette = 'Reds', direction = 1)
# direction ensures that darker = higher values
```

If 25 were a sensible "middle point", then a diverging color would make more sense.

```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.355, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  ) +
  coord_cartesian(expand = FALSE) +
  scale_fill_distiller(type = 'div')
```


To fix the "midpoint" to a certain value, you could also use `scale_fill_gradient2"
Colors can be found on https://colorbrewer2.org/#type=diverging&scheme=BrBG&n=5


```{r}
standard_heatmap +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.355, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  ) +
  coord_cartesian(expand = FALSE) +
  scale_fill_gradient2 (
    low = '#018571',
    mid = "#f5f5f5",
    high = '#a6611a',
    midpoint = 30,
  )
```

Add values to cells.
Remember to move texts and set color dependent on `generated_data$n`

```{r}
standard_heatmap +
  geom_text(
    aes(label = n),
    color = if_else(generated_data$n > 30, 'white', 'black'),
    size = 5.5,
    hjust = 1,
    nudge_x = 4.25,
    nudge_y = -0.3,
    family = 'Source Sans Pro'
  ) +
  theme(
    plot.title.position = 'plot',
    plot.subtitle = element_text(margin = margin(b = 3.25, unit = 'cm')),
    legend.position = c(0.355, 1.125),
    legend.direction = 'horizontal'
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = unit(20, 'cm'),
      barheight = unit(0.5, 'cm'),
      title.position = 'top'
    )
  ) +
  coord_cartesian(expand = FALSE) +
  scale_fill_distiller(type = 'seq', palette = 'Reds', direction = 1)
```




