---
theme: ../theme.scss
format: 
  html: 
    fig-height: 5
    fig-dpi: 300
    fig-width: 8.88
    fig-align: center
knitr: 
  opts_chunk: 
    collapse: true
---

```{r}
setwd(here::here('90_Aug_21_2024'))
library(tidyverse)

theme_set(
  theme_minimal(
    base_size = 16,
    base_family = 'Source Sans Pro'
  ) +
    theme(panel.grid.minor = element_blank())
) 
```


Many greetings from Munich, Germany. Are you done with tables yet? I hope not. The R ecosystem has more to offer.

Over the last two weeks, we have covered the fantastic `{gt}` and `{reactable}` packages.
Both are excellent at what they do.
`{gt}` has an impeccable workflow to create static tables and `{reactable}` is the ultra-champ at creating interactive tables.
Today, we get to know another champion.
Namely, `{flextable}`.
But before that, time for two short little announcement:


---

# `{reactable}` video tutorial

Looking for a video version of last week's newsletter?
Well, look no further.
I've created a video intro on `{reactable}` with a couple more features that we didn't even cover in the newsletter. Check it out on Youtube:

LINK


# Data Cleaning Master Class Part 2

Part 2 of my [**Data Cleaning Master Class**](https://data-cleaning.albert-rapp.de/) is about to be released.
I'm in the midst of final video editing and then everything Part 2 will be released within the next 14 days.

Part 2 will cover 

- a myriad of data formats, 
- show you how to read them into R and 
- clean them so that you can work with that data.

And the most important part:
As I release another part of the course, the price of the course will rise to $175.
So if you've been on the fence and want to save some money, now is your chance:

LINK BOX

Now, let's dive into this week's issue.

---

# What can `{flextable}` do for me?

The biggest feature of `{flextable}` is that it has the strongest export capabilities to Microsoft Office files like Word documents or Powerpoint slides.
Thus, if you frequently find yourself copying a data table over to Word or Powerpoint, then look no more.
Let `{flextable}` handle that for you.

---

# Mini Tutorial

## Fake data

Let’s create a similar table that we have created in the last two weeks.
By now, you probably now the data set by heart.


```{r}
#| warning: false
#| message: false
library(tidyverse)
hawaiian_sales <- gt::pizzaplace |> 
  filter(name == 'hawaiian') |> 
  mutate(
    month = month(
      date, label = TRUE, abbr = FALSE,
      locale = 'en_US.UTF-8' # English month names
    ),
    quarter = paste0('Q', quarter(date))
  ) |> 
  summarise(
    sales = n(),
    revenue = sum(price),
    .by = c(month, quarter)
  )
hawaiian_sales
```


## Base Layer

Just like all the other table packages, the first step is really simple with `{flextable}`.
Just throw your data into the `flextable()` function.

```{r}
#| warning: false
#| message: false
library(flextable)
flextable(hawaiian_sales)
```

TABLE


---

## Use better column names

Now, let's do better column names again.
Let's go for title case.
And the nice thing with `{flextable}` is that it allows a pipe-able code style like `{gt}`.

```{r}
flextable(hawaiian_sales) |> 
  set_header_labels(
    month = 'Month',
    quarter = 'Quarter',
    sales = 'Sales',
    revenue = 'Revenue'
  )
```

TABLE

---

## Title & Subtitle

For adding a nice title and subtitle to your plot, you can add so-called header lines.

```{r}
flextable(hawaiian_sales) |> 
  set_header_labels(
    month = 'Month',
    quarter = 'Quarter',
    sales = 'Sales',
    revenue = 'Revenue'
  ) |> 
  add_header_lines(
    'Hawaiian Pizza Sales in 2015'
  ) |> 
  add_header_lines(
    'Based on the fake `pizzaplace` data from `{gt}`'
  )
```

TABLE

---

## Format numbers

The numbers in the revenue column correspond to dollar amounts. We can format them by specifying a column format with one of the `colformat_*()` functions.
Unfortunately, there's no such thing as `colformat_currency()` so we'll have to improvise with `colformat_double()`.

```{r}
flextable(hawaiian_sales) |> 
  set_header_labels(
    month = 'Month',
    quarter = 'Quarter',
    sales = 'Sales',
    revenue = 'Revenue'
  ) |> 
  add_header_lines(
    'Hawaiian Pizza Sales in 2015'
  ) |> 
  add_header_lines(
    'Based on the fake `pizzaplace` data from `{gt}`'
  ) |> 
  colformat_double(
    j = 'revenue',
    digits = 2,
    prefix = '$'
  )
```

TABLE

---

## Add groups

Now, I want to structure my tables into quarters. 
The easiest way to do that is to use the quarter column in our data set for grouping.

But `flextable()` doesn't allow to do that inside of the function.
Instead one has to pass in the data as grouped data set.


```{r}
hawaiian_sales |> 
  as_grouped_data(groups = 'quarter') |> 
  flextable() |> 
  set_header_labels(
    month = 'Month',
    quarter = 'Quarter',
    sales = 'Sales',
    revenue = 'Revenue'
  ) |> 
  add_header_lines(
    'Hawaiian Pizza Sales in 2015'
  ) |> 
  add_header_lines(
    'Based on the fake `pizzaplace` data from `{gt}`'
  ) |> 
  colformat_double(
    j = 'revenue',
    digits = 2,
    prefix = '$'
  )
```

TABLE

---

## Add summaries

For adding summaries to a table, `{flexbox}` is actually one of the trickiest packages.
You will have to calculate totals yourself and assemble a new data set.

```{r}
order_of_months_w_total <- c(
  levels(hawaiian_sales$month),
  'Total'
)

totals <- hawaiian_sales |> 
  summarise(
    month = factor(
      'Total',
      levels = order_of_months_w_total
    ),
    sales = n(),
    revenue = sum(revenue),
    .by = quarter
  ) 

combined_data <- hawaiian_sales |> 
  mutate(
    month = factor(
      month, 
      levels = order_of_months_w_total,
      ordered = FALSE
    )
  ) |> 
  bind_rows(totals) |> 
  arrange(quarter, month)
combined_data
```

And then you can throw that into the `{flextable}` again.

```{r}
table_with_totals <- combined_data |> 
  as_grouped_data(groups = 'quarter') |> 
  flextable() |> 
  set_header_labels(
    month = 'Month',
    quarter = 'Quarter',
    sales = 'Sales',
    revenue = 'Revenue'
  ) |> 
  add_header_lines(
    'Hawaiian Pizza Sales in 2015'
  ) |> 
  add_header_lines(
    'Based on the fake `pizzaplace` data from `{gt}`'
  ) |> 
  colformat_double(
    j = 'revenue',
    digits = 2,
    prefix = '$'
  )
table_with_totals
```


TABLE


## Change row styling

Finally, to add a little bit more style and visual structure let us make the group rows blue.
And the totals bold.
For that, you can use the `bg()` and `bold()` functions to apply the styles.
And to target specific rows, you need to see `i` to a function (using the `~` notation) that evaluates to `TRUE`s and `FALSE`s.

```{r}
styled_table <- table_with_totals |> 
  bg(
    i = ~!is.na(quarter),
    bg = '#E7EDF3'
  ) |> 
  bold(
    i = ~(month == 'Total')
  )
styled_table
```

TABLE


---

# Export

Finally, we can make use of `{flexbox}` fantastic export capabilities and export to Word....

```{r}
styled_table |> 
  save_as_docx(
    path = 'word_export.docx'
  )
```

IMAGE

...or to Powerpoint (though you might want to do that with a table that fits into a rectangular slide)

```{r}
styled_table |> 
  save_as_pptx(
    path = 'powerpoint_export.pptx'
  )
```

IMAGE 2


