---
theme: ../theme.scss
format: 
  html: 
    fig-height: 5
    fig-dpi: 300
    fig-width: 8.88
    fig-align: center
knitr: 
  opts_chunk: 
    collapse: true
editor_options: 
  chunk_output_type: console
---

```{r}
setwd(here::here('94_Oct_02_2024'))
library(tidyverse)

theme_set(
  theme_minimal(
    base_size = 16,
    base_family = 'Source Sans Pro'
  ) +
    theme(panel.grid.minor = element_blank())
) 
```

Last week, I told you about the `{officer}` package that let's you build a Word document from scratch using R.
But that was a bit much wasn't it?
Clearly, no one wants to create a WHOLE Word document from R.
It's much easier to use Word to create a document template and then use the `{officer}` package to insert data-related things.
So today, I'm showing you how to do that.

---

# Find a template

First, let us download a template that we want to use.
The cool thing about the `{officer}` package is that the look of the template doesn't matter.
We can even use one that uses fancy visual elements and styles.
Like this one here:

IMG

---

# Insert placeholder

Now that we have that template, we can modify it to our needs.
For example, I want to show you how to create this Word document from the template:

FINAL IMG

So, a couple of quick modifications here and there and ready is our template:

TEMPLATE IMG

---

# Insert a date

As you can see, our template contains a few placeholders like "[DATE]".
These can be replaced using functions from the `{officer}` package.
The key function to replace text is `body_replace_all_text()`.

This function works by looking for a specific string and then replacing it with something else.
Let's try that with the "[DATE]" string.

```{r}
#| warning: false
#| message: false
library(officer)
doc_template <- read_docx('word_templates/letter.docx') 
doc_template |> 
  body_replace_all_text(
    old_value = '[DATE]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  print(target = '00_date_replacement.docx')
```

CODE

IMG

---

# Watch out! Regular expressions

Oh no.
This doesn't look right.
Something weird is going on.
And the weirdness has a name: It's regular expressions.

You see, the thing that we stick into the `old_value` argument is a regular expression.
This means it uses a complex but powerful language to describe patterns (btw, my [data cleaning masterclass](https://data-cleaning.albert-rapp.de/) teaches that powerful language.) 

In this language the brackets `[` and `]` have a special meaning.
And that meaning is what destroys our text replacement.
But we can tell that regular expression to use the brackets as is using `\\`.
That way, we get our desired result.

```{r}
doc_template <- read_docx('word_templates/letter.docx') 
doc_template |> 
  body_replace_all_text(
    old_value = '\\[DATE\\]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  print(target = '01_date_replacement_regex.docx')
```

CODE

IMG

---

# Insert an image


Nice! 
This worked pretty well.
Now let's try something harder.
Let's insert a company logo.

For that I have created a generic logo.
It looks like this:

IMG

Unfortunately, there is no function to replace a text with an image.
There is only `body_add_img()`.

So in order to replace the "[COMPANY_LOGO]" placeholder with an image, we have to do three things:

1. Navigate a fictitious cursor to the position of the paragraph that contains "COMPANY_LOGO"

2. Remove the paragraph (as it only contains the placeholder)

3. Add the image into that paragraph

And for all three of thoses steps there are three functions that do exactly that.

```{r}
doc_template <- read_docx('word_templates/letter.docx') 
doc_template |> 
  cursor_reach('COMPANY_LOGO') |> 
  body_remove() |> 
  body_add_img(
    src = 'company_logo.png',
    width = 0.75 * 1201 / 376,
    height = 0.75,
    pos = 'before' # placement relative to cursor pos
  ) |> 
  body_replace_all_text(
    old_value = '\\[DATE\\]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  print(target = '02_img_replacement.docx')
```

CODE

IMG

---

# Right-align the image

Unfortunately, this doesn't look nice.
We want to have our image right-aligned.
That's why we have to make sure that the image is inserted in a right-aligned paragraph.

This can be done via the `style` argument in `body_add_img()`.
But this argument requires the name of a pre-defined style.
Hence, we have to define such a style using `docx_set_paragraph_style()` and the `fp_par()` function from last time.

```{r}
doc_template <- read_docx('word_templates/letter.docx') 

# Add style to document
doc_template <- doc_template |> 
  docx_set_paragraph_style(
    style_id = 'right-aligned-img',
    style_name = 'Right aligned Image',
    fp_p = fp_par(text.align = 'right')
  )

doc_template |> 
  cursor_reach('COMPANY_LOGO') |> 
  body_remove() |> 
  body_add_img(
    src = 'company_logo.png',
    width = 0.75 * 1201 / 376,
    height = 0.75,
    style = 'Right aligned Image', # Use here
    pos = 'before'
  ) |> 
  officer::body_replace_all_text(
    old_value = '\\[DATE\\]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  print(target = '03_img_replacement_right_aligned.docx')
```

CODE

And with that, we do have a nice right-aligned company logo.

IMG

---

# Insert line breaks

Unfortunately, our date seems to have been moved by that.
No problem.
We just have to define the image paragraph style to have more padding at the bottom.


```{r}
doc_template <- read_docx('word_templates/letter.docx') 
doc_template_with_style <- doc_template |> 
  docx_set_paragraph_style(
    style_id = 'rightalignedimg',
    style_name = 'Right aligned Image',
    fp_p = fp_par(
      text.align = 'right', 
      padding.bottom = 100 ## Add padding bottom
    )
  )
```

CODE

And once we have done that, we can reexecute the same code as before.
Et voilà, we have a nicer date position.

IMG


```{r}
doc_template_with_style |> 
  cursor_reach('COMPANY_LOGO') |> 
  body_remove() |> 
  body_add_img(
    src = 'company_logo.png',
    width = 0.75 * 1201 / 376,
    height = 0.75,
    style = 'Right aligned Image',
    pos = 'before'
  ) |> 
  body_replace_all_text(
    old_value = '\\[DATE\\]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  print(target = '04_img_and_date_replacement.docx')
```


---

# Replace remaining placeholders

Next, let us replace all of the remaining placeholders as well.
We use `body_add_img()` just like before.
And just like it was the case with the date replacement, the style in which we formatted the placeholders always remains the same (even if they're hideous.)

```{r}
#| echo: false
doc_template <- read_docx('word_templates/letter.docx') 
doc_template_with_style <- doc_template |> 
  docx_set_paragraph_style(
    style_id = 'rightalignedimg',
    style_name = 'Right aligned Image',
    fp_p = fp_par(
      text.align = 'right', 
      padding.bottom = 100 ## Add padding bottom
    )
  )
```

```{r}
# Re-compute `doc_template_with_style`
# Otherwise company logo already replaced
fully_replaced_doc <- doc_template_with_style |> 
  cursor_reach('COMPANY_LOGO') |> 
  body_remove() |> 
  body_add_img(
    src = 'company_logo.png',
    width = 0.75 * 1201 / 376,
    height = 0.75,
    style = 'Right aligned Image',
    pos = 'before'
  ) |> 
  body_replace_all_text(
    old_value = '\\[DATE\\]',
    new_value = lubridate::now() |> format('%d.%m.%Y')
  ) |> 
  ## REMAINING REPLACEMENTS
  body_replace_all_text(
    old_value = 'RECIPIENT',
    new_value = 'Reader'
  ) |>
  body_replace_all_text(
    old_value = '\\[PURPOSE\\]',
    new_value = 'demonstrate that you can fill Word docs using R'
  ) |> 
  body_replace_all_text(
    old_value = '\\[ADJECTIVE\\]',
    new_value = 'STUPID'
  ) |> 
  body_replace_all_text(
    old_value = '\\[SENDER_NAME\\]',
    new_value = 'Albert'
  ) |> 
  body_replace_all_text(
    old_value = '\\[SENDER_URL\\]',
    new_value = 'https://3mw.albert-rapp.de/'
  ) 

fully_replaced_doc |> 
  print(target = '05_all_placeholders_replaced.docx')
```

CODE

IMG

---

# Insert ggplot

And to round it all off, let me show you how to insert a ggplot before the greetings.
All you have to do is move the fictitious cursor again. 
This time, we move the cursor to to the greetings and then insert the plot with `body_add_gg()`.

```{r}
fully_replaced_doc |> 
  cursor_reach('Warm Regards') |>
  body_add_gg(
    pos = 'before',
    height = 3,
    width = 3 * 16 / 9,
    value = {
      palmerpenguins::penguins |> 
        ggplot(aes(flipper_length_mm, bill_length_mm)) +
        geom_point(
          aes(fill = species), 
          size = 3, 
          shape = 21, 
          col = 'black',
          alpha = 0.9
        )
    }
  ) |> 
  print(target = '06_with_ggplot.docx')
```

CODE

IMG
















