
Basic plot

```{r}
library(tidyverse)
penguins <- palmerpenguins::penguins |> 
  filter(!is.na(sex))

basic_plot <- penguins |> 
  ggplot(aes(x = bill_length_mm, y = body_mass_g, fill = species)) +
  geom_point(shape = 21, size = 5, alpha = 0.85, color = 'grey10') +
  labs(
    x = 'Bill length (in mm)',
    y = element_blank(),
    fill = 'Species',
    title = 'Penguins from the Palmer Archipelago',
    subtitle = 'Penguin weights (in g)',
    caption = 'Data: {palmerpenguins} R package'
  )
basic_plot
```

Add theme_minimal

```{r}
basic_plot +
  theme_minimal()
```

Increase font size and set a nice base family

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro')
```

Change colors e.g. Okabe Ito

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6))
```

Or using colors from [Coolors](https://coolors.co/generate)

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = c('#da2c38', '#FED18C', '#30C5FF'))
```

Pro tip: Fix colors to species via names

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(
    values = c(
      'Adelie'= '#da2c38', 
      'Chinstrap'= '#FED18C', 
      'Gentoo'= '#30C5FF'
    )
  )
```

Get rid of too many grid lines

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank()
  )
```

Make lines less noticeable

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    )
  )
```

Move plot titles to left

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot'
  )
```

Move caption to right

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot'
  )
```


Make title flashy, i.e. larger and different font.
Pro tip: Use `rel()` to make size relative to base size.

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7)
    )
  )
```

Give some room to y-axis label

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7),
      margin = margin(b = 7, unit = 'mm')
    )
  )
```

Change subtitle text

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7),
      margin = margin(b = 7, unit = 'mm')
    ),
    plot.subtitle = element_text(size = rel(1.1))
  )
```


Optional: Make texts lighter.

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7),
      margin = margin(b = 7, unit = 'mm')
    ),
    plot.subtitle = element_text(size = rel(1.1)),
    text = element_text(color = 'grey20'),
    axis.text = element_text(color = 'grey30'),
  )
```

Optional: Add framing to panel

```{r}
basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7),
      margin = margin(b = 7, unit = 'mm')
    ),
    plot.subtitle = element_text(size = rel(1.1)),
    text = element_text(color = 'grey20'),
    axis.text = element_text(color = 'grey30'),
    panel.background = element_rect(color = 'grey90')
  )
```


Move to a function

```{r}
my_base_theme <- function() {
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      color = 'grey90', linetype = 2
    ),
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    plot.title = element_text(
      family = 'Merriweather', 
      size = rel(1.7),
      margin = margin(b = 7, unit = 'mm')
    ),
    plot.subtitle = element_text(size = rel(1.1)),
    text = element_text(color = 'grey20'),
    axis.text = element_text(color = 'grey30'),
    panel.background = element_rect(color = 'grey90')
  )
}

basic_plot +
  theme_minimal(base_size = 16, base_family = 'Source Sans Pro') +
  scale_fill_manual(values = thematic::okabe_ito(6)) +
  my_base_theme()
```
